---
title: "Trust to local authority mapping overview"
author: "Sophie Meakin"
output: html_document
vignette: >
  %\VignetteIndexEntry{Trust to local authority mapping overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

# Summary

This package contains two many-to-many mappings between local authority districts (UTLA and LTLA) and NHS Acute Trusts in England:

* Trust-UTLA mapping: `covid19.nhs.data::trust_utla_mapping`
* Trust-LTLA mapping: `covid19.nhs.data::trust_ltla_mapping`

These mappings each contain the following variables:

* `trust_code`: the three-digit organisation code for NHS Trusts.
* `geo_code`: the nine-digit identifier for local authorities (UTLA or LTLA) in England
* `p_trust`: the estimated proportion of admissions to a given Trust were admitted from a given local authority (UTLA or LTLA)
* `p_geo`: the estimated proportion of admissions from a given local authority (UTLA or LTLA) that were admitted to a given Trust.

These mappings are probabilistic estimates based on total COVID-19 discharges until 30 September. These mappings can be used to estimate COVID-19 hospital admissions at the local authority level, or to estimate "relevant" community COVID-19 cases for an NHS Trust. These are estimates only and may not accurately reflect the truth; see limitations below.


# Methods

The mappings in this package are derived from a raw mapping provided by NHS England, based on the Secondary Uses Service (SUS) healthcare data for England. This raw mapping counts the number of COVID-19 hospital discharges of between 01 January 2020 and 30 September 2020 from an NHS site (hospital) to lower-tier local authorities (LTLAs).

The mappings in this package are summaries of the raw mapping. Site-LTLA pairs are aggregated to Trust-LTLA and Trust-UTLA pairs and to maintain anonymity we excluded any Trust-UTLA or Trust-LTLA pairs where the number of discharges is less than 10. Finally, we report only the proportion of admissions by Trust (`p_trust`) or local authority (`p_geo`).

The mapping includes the following Trust mergers that have taken place since January 2020:

```{r mergers}
library(readr)
library(dplyr)
library(knitr)

read_csv("https://raw.githubusercontent.com/epiforecasts/covid19.nhs.data/main/data-raw/raw/england_trusts/trust_mergers.csv") %>%
  select(`Trust code` = trust_code_old, `New Trust code` = trust_code_new, `Merger date` = from_date) %>%
  kable()

```

The mapping also includes the creation of Buckinghamshire unitary authority (UTLA and LTLA; E06000060), replacing Buckinghamshire county (UTLA; E10000002) and four lower-tier local authorities: Aylesbury Vale (E07000004), Chiltern (E07000005), South Bucks (E07000005) and Wycombe (E07000007).

See [vignettes/create_public_mapping](https://epiforecasts.io/covid19.nhs.data/articles/create_public_mapping.html) for full details of how these public mappings were created.


# Usage

The mapping is provided as a built in package dataset and can be loaded directly using `covid19.nhs.data::trust_utla_mapping` and `covid19.nhs.data::trust_ltla_mapping` or as shown below.

## Quick start

Load the library with: 

```{r package, echo = TRUE, eval = TRUE}
library(covid19.nhs.data)
```

Load the basic Trust-UTLA mapping with `trust_utla_mapping`:

```{r load_raw_mapping, echo = TRUE, eval = TRUE}

trust_utla_mapping %>%
  head() %>%
  knitr::kable()

```

Add Trust names and local authority names to the raw mapping with `get_names()`:

```{r add_names, echo = TRUE, eval = TRUE}

trust_utla_mapping %>% 
  get_names() %>%
  head() %>%
  knitr::kable()

```

Summarise the mapping for a given Trust (as defined by `trust`, a Trust code) with `summarise_mapping()`; this will return a table (under `summary_table`) and a visualisation of the Trust-UTLA mapping (under `summary_plot`):

```{r summarise for trust, echo = TRUE, eval = TRUE}

summary <- summarise_mapping(mapping = trust_utla_mapping, trust = "R0A")

summary$summary_table %>%
  knitr::kable()

summary$summary_plot

```

Summarise the mapping for a given local authority (as defined `geography`, either a UTLA or LTLA code) with `summarise_mapping()`; this will return a table (under `summary_table`):

```{r summarise for ltla, echo = TRUE, eval = TRUE}

summary <- summarise_mapping(mapping = trust_utla_mapping, geography = "E09000012")

summary$summary_table %>%
  knitr::kable()

```


