---
title: "Mapping local authorities to NHS Acute Trusts in England"
author: "Sophie Meakin"
output: html_document
vignette: >
  %\VignetteIndexEntry{Mapping local authorities to NHS Acute Trusts in England}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(knitr)
```

# Summary

This package contains many-to-many mappings between local authority districts (UTLA and LTLA) and NHS Acute Trusts in England, based on COVID-19 hospital admissions:

* Trust-UTLA mapping: `covid19.nhs.data::trust_utla_mapping`
* Trust-LTLA mapping: `covid19.nhs.data::trust_ltla_mapping`

These mappings can be used to map data reported by local authority (e.g. testing) to NHS Trusts, or to map Trust-level data (e.g. admissions) to local authorities.

The mappings contain the following variables:

* `trust_code`: the three-digit organisation code for NHS Trusts.
* `geo_code`: the nine-digit identifier for local authorities (UTLA or LTLA) in England
* `p_trust`: the proportion of admissions to a given Trust were admitted from a given local authority (UTLA or LTLA)
* `p_geo`: the proportion of admissions from a given local authority (UTLA or LTLA) that were admitted to a given Trust.

# Usage

The mapping is provided as a built in package dataset and can be loaded directly using `covid19.nhs.data::trust_utla_mapping` and `covid19.nhs.data::trust_ltla_mapping` or as shown below.


## Quick start

Load the library with: 

```{r package, echo = TRUE, eval = TRUE}
library(covid19.nhs.data)
```

Load the basic Trust-UTLA mapping with `trust_utla_mapping`:

```{r load_raw_mapping, echo = TRUE, eval = TRUE}

trust_utla_mapping %>%
  head() %>%
  knitr::kable()

```

Add Trust names and local authority names to the raw mapping with `get_names()`:

```{r add_names, echo = TRUE, eval = TRUE}

trust_utla_mapping %>% 
  get_names() %>%
  head() %>%
  knitr::kable()

```

Summarise the mapping for a given Trust (as defined by the Trust code) with `summarise_mapping()`; this will return a table (under `summary_table`) and a visualisation of the Trust-UTLA mapping (under `summary_plot`):

```{r summarise for trust, echo = TRUE, eval = TRUE}

summary <- summarise_mapping(mapping = trust_utla_mapping, trust = "R0A")

summary$summary_table %>%
  knitr::kable()

summary$summary_plot

```

Summarise the mapping for a given local authority (as defined by the UTLA or LTLA code) with `summarise_mapping()`; this will return a table (under `summary_table`):

```{r summarise for ltla, echo = TRUE, eval = TRUE}

summary <- summarise_mapping(mapping = trust_utla_mapping, geography = "E09000012")

summary$summary_table %>%
  knitr::kable()

```



# Methods

The raw (unpublished) mapping was provided by NHS England, This mapping provides the number of COVID-19 spells (discharges) between 01 January 2020 and 30 September 2020 by hospital-LTLA pairs, based on the Secondary Uses Service (SUS) healthcare data for England.

We have summarised the raw mapping between Trusts and local authorities (UTLA and LTLA), and to maintain anonymity we excluded any Trust-UTLA or Trust-LTLA pairs where the number of spells is less than 10. Finally, we report only the proportion of admissions by Trust (`p_trust`) or local authority (`p_geo`).


