---
title: "Overview of LTLA- and UTLA-Trust COVID-19 mappings"
author: "Sophie Meakin"
output: html_document
vignette: >
  %\VignetteIndexEntry{Mapping overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

pacman::p_load(
  magrittr,
  dplyr,
  knitr,
  kableExtra
)

```

# Summary

The `covid19.nhs.data` package contains four many-to-many (overlapping) mappings between local authority districts (upper- and lower-tier) and NHS Acute Trusts in England. Each mapping contains the following variables:

-   `geo_code`: A 9-digit identifier for local authorities (UTLA or LTLA) in England.
-   `trust_code`: A 3-digit organisation code for NHS Trusts.
-   `p_geo`: The proportion of all admissions from a given local authority (LTLA or UTLA) that were admitted to a given Trust, estimated from one of two datasets (see below).
-   `p_trust`: The proportion of all admissions to a given Trust that were admitted from a given local authority (LTLA or UTLA), estimated from one of two datasets (see below).

These mappings are probabilistic estimates based on Secondary Uses Service (SUS) healthcare data for England (January - September 2020, inclusive), and linked COVID-19 cases and admissions (June 2020 - May 2021, inclusive). Please see the vignette ["Creating the public mappings"](https://epiforecasts.io/covid19.nhs.data/articles/create_public_mapping.html) for details of how the mappings are made.

These mappings can be used to estimate COVID-19 hospital admissions at the local authority level, or to estimate "relevant" community COVID-19 cases for an NHS Trust. These are estimates only and may not accurately reflect the truth; see limitations below.

# Usage

## Get started

Install the package from GitHub:

```{r, echo = TRUE, eval = FALSE}
devtools::install_github("epiforecasts/covid19.nhs.data")
```

and load the library:

```{r package, echo = TRUE, eval = TRUE}
library(covid19.nhs.data)
```

## Access the mappings

Access the mappings with `load_mapping()`, specifying the geographical scale ("ltla" or "utla") and the data source ("sus" or "link"):

```{r echo = TRUE, eval = TRUE}

load_mapping(scale = "ltla", source = "link") %>%
  head() %>%
  kable() %>%
  kable_styling()

```

Add Trust names and local authority names to the raw mapping with `get_names()`:

```{r add_names, echo = TRUE, eval = TRUE}

load_mapping(scale = "ltla", source = "link") %>%
  get_names(geo_names = ltla_names) %>%
  head() %>%
  kable() %>%
  kable_styling()

```

Summarise the mapping with `summarise_mapping()`.
For an NHS Trust (specified by the argument `trust`), this will return a table and a visualisation, both describing `p_trust`:

```{r summarise for trust, echo = TRUE, eval = TRUE}

mapping <- load_mapping(scale = "ltla", source = "link")
summary <- summarise_mapping(trust = "RYR",
                             mapping = mapping,
                             shapefile = england_ltla_shape,
                             geo_names = ltla_names)

summary$summary_table %>%
  kable() %>%
  kable_styling()

summary$summary_plot

```

For a local authority (specified by the argument `geography`), this will return a table describing `p_trust`:

```{r summarise for ltla, echo = TRUE, eval = TRUE}

summary <- summarise_mapping(geography = "E09000012",
                             mapping = mapping,
                             shapefile = england_ltla_shape,
                             geo_names = ltla_names)

summary$summary_table %>%
  kable() %>%
  kable_styling()

```
